---
title: "Initial Korean covid 19"
author: "Aeilani Felix "
date: "Spring 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
 
## Methods
  
  We will analyze the Sex of the patient and the result  if they are alive or dead. I will use the variables in any collection of cat~cat.

# Your Question - Is dying independent of sex?

## Graphical results - use ggplot commands to do this  
```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

  The data in this plot shows us the results of the people that are alive and dead. We can see that Twice as many males died than females.The disease is killing more males than females according to the data from Korea.

## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

This shows us all of the people that are still alive, 56 are females and 36 are males. The number of dead people were 44 females and 64 males. The percentage was 98.7% of females remained alive and only 1.4% females died and 97% of males remained alive while 3% are dead. 

## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The test and the fisher test highlights the disparity between the sexes of male and female in regards to being alive or dead. The p-value of 0.0009 confirms that there is a significant difference in the data with males about 2 times more likely than females to die from covid at least in the case of Korea.

## Conclusions

In summary, Our proportion-based bar charts shows a variation in proportion among those who died favoring men, and this furthered by our numerical results which provided the exact percentage to fortify that claim. Finally our inferential results provided the exact counts of the prior tests, but also yielded a low p-value, helping to certify that our results had not purely occurred through by chance. Our odds ratio 2.3 was the single most compelling piece of data generated, helping to reinforce the alternative hypothesis that dying is actually dependent at least in part on sex.  
